---
title: "Called Strike Package"
author: "Jim Albert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Called Strike Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

The ```CalledStrike``` package (version 0.5.2)  provides several visualizations of smoothed measures over the zone using Statcast data.  

### Producing the Graphs

Here are the basic steps to produce these graphs.

-  By use of one of the set_up functions, define the relevant dataset and define any new variables if needed.  

- Use a generalized additive model to fit a smooth model to the measure over the (plate_x, plate_z) surface.  In the case where the response is binary (1 or 0), fit a binary gam with logit link.  When the response is continuous, such as a launch velocity, use a continuous-response gam.

- Use the model estimates to predict the measure over a 50 by 50 grid over the zone.  (The ```grid_predict()``` function is used.)

- Use the ```geom_tile()``` or ```geom_contour_fill()``` geometric objects to graph the predicted values


### Graph Types

There are 13 possible measures on the zone.  For each type of measure, one can construct a tile plot or a filled contour plot.  For a filled contour plot, one has the option of specifying a vector of contour line values by use of the ```L``` argument.  For all graphs, one has the option of specifying a title by use of the ```title``` argument.

Graph Type | Tile Plot Function | Contour Plot Function
---- | ---- | ----
called strikes | ```called_strike_plot()``` | ```called_strike_contour()```
swing | ```swing_plot()``` | ```swing_contour()```
contact on swing | ```contact_swing_plot()``` | ```contact_swing_contour()```
miss on swing | ```miss_swing_plot()``` | ```miss_swing_contour()```
in-play on swing | ```inplay_swing_plot()``` | ```inplay_swing_contour()```
launch speed | ```ls_plot()``` | ```ls_contour()```
launch angle | ```la_plot()``` | ```la_contour()```
spray angle | ```sa_plot()``` | ```sa_contour()```
batting average | ```hit_plot()``` | ```hit_contour()```
home run | ```home_run_plot()``` | ```home_run_contour()```
wOBA | ```woba_plot()``` | ```woba_contour()```
expected batting average | ```ehit_plot()``` | ```ehit_contour()```
expected wOBA | ```ewoba_plot()``` | ```ewoba_contour()```

### Input is a Data Frame or a List

The input to each function is a Statcast data frame or a list of Statcast data frames.  If the input is a list of data frames, one will see a paneled graphical display which is useful for comparison.

### Example:  Nolan Arenado

The package contains the dataset ```sc_sample``` containing Statcast data for all swings for 10 hitters in the 2019 season.

First load in the ```CalledStrike``` package and the ```dplyr``` package which will be used for the ```filter()``` function.

```{r}
library(CalledStrike)
library(dplyr)
```

One of the hitters is Nolan Arenado and I will collect Arenado's data for this demonstration.

```{r}
na <- filter(sc_sample,
             player_name == "Nolan Arenado")
```

I'll look at locations of called strikes.

```{r}
called_strike_contour(na, L = c(0.25, .5, .75))
```

I'll look where Arenado swung at the pitch.

```{r}
swing_contour(na, L = seq(0, 1, by = 0.05))
```

Did he make contact with the pitch?

```{r}
contact_swing_contour(na, L = seq(0, 1, by = 0.05))
```

For the balls where he made contact, we look at the launch velocity.

```{r}
ls_contour(na, L = seq(60, 100, by = 2))
```

What was the expected batting average on the balls in play?

```{r}
ehit_contour(na, L = seq(0, .5, by = 0.01))
```

What was the expected wOBA on balls in play?

```{r}
ewoba_contour(na, L = seq(0, .5, by = 0.01))
```

### Example: Comparison Graphs

Suppose we are interested in comparing Manny Machado with Matt Chapman with respect to launch velocity.

```{r}
mm <- filter(sc_sample,
             player_name == "Manny Machado")
mc <- filter(sc_sample,
             player_name == "Matt Chapman")
two_players <- list(mm, mc)
names(two_players) <- c("Machado", "Chapman")
ls_contour(two_players, L = seq(60, 100, by = 2),
           title = "Launch Velocities of 2 Hitters")
```
