---
title: "CalledStrike"
author: "Jim Albert"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

Load the package.  It assumes that the packages ggplot2, dplyr, baseballr, and mgcv have already been loaded.

```{r}
library(CalledStrike)
```

#### Collect Data

Here we find the MLBAM id for Mike Trout.  The get_sc_data() function (wrapper for the baseballr function) will download the Statcast data for Trout for the 2018 season.

```{r}
mt_id <- get_id("Mike Trout")
mt <- get_sc_data(mt_id, season = 2018)
```

#### Called Strikes

This constructs a tile plot for the probability that a called pitch is a strike.

```{r}
called_strike_plot(mt, "Mike Trout Called Strikes")
```

#### Miss Swing Rates

Constructs a tile plot for the probability that Trout misses on a swing.

```{r}
miss_swing_plot(mt, "Mike Trout Miss Swing Rates")
```

#### Zones of Hits and Home Runs

Constructs a tile plot for the probability that Trout hits a home run on a ball in-play.

```{r}
home_run_plot(mt, "Mike Trout Home Run Rates")
```

Constructs a tile plot for the probability that Trout gets a hit on a ball in-play.

```{r}
hit_plot(mt, "Mike Trout Hit Rates")
```

#### Statcast Variables

Constructs a tile plot for the launch speed on a ball in-play.

```{r}
ls_plot(mt, "Mike Trout Launch Speeds")
```

Constructs a tile plot for the launch angle on a ball in-play.

```{r}
la_plot(mt, "Mike Trout Launch Angles")
```

Constructs a tile plot for the spray angle on a ball in-play.

```{r}
sa_plot(mt, "Mike Trout Spray Angles")
```

#### Contour Plots

The wrapper functions above all use tile plots, but these can be drawn using contour plots where one inputs the value for the contour lines.

Here is a contour plot version of Trout's in-play hit rates.

```{r}
mt %>%
    setup_inplay() %>%
    hr_h_gam_fit(HR = FALSE) %>%
    contour_plot_p(P = c(0.3, 0.4, 0.5)) +
    ggtitle("Mike Trout Hit Rates") +
    centertitle()
```

Here is a contour plot version of Trout's in-play launch speeds.

```{r}
mt %>%
    setup_inplay() %>%
    ls_gam_fit() %>%
    contour_plot_m(M = c(80, 90, 100)) +
    ggtitle("Mike Trout Launch Speeds") +
    centertitle()
```

#### Comparison Plots

Let's compare Trout's performance for the 2017 and 2018 seasons. 

First collect data for the two seasons:

```{r}
mt_07 <- get_sc_data(mt_id, season = 2017)
mt_08 <- get_sc_data(mt_id, season = 2018)
```

Compare launch speeds for two seasons:

```{r}
compare_ip(mt_07, mt_08,
           "2017 Trout", "2018 Trout", 
           type = "ls")
```


Compare home run probabilities for the two seasons:

```{r}
compare_ip(mt_07, mt_08,
           "2017 Trout", "2018 Trout", 
           type = "hr")
```