---
title: "Plot Location Graphs"
author: "Jim Albert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot Location Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(CalledStrike)
library(dplyr)
```

#### Introduction

There are three functions for visualizing pitch locations.

- The function ```location_compare()``` will graph the pitch location for a data frame or list of data frames.

- The function ```location_count()``` will show the locations of pitches for a specific pitcher on a particular count.

- The function ```location_count_compare()``` will graph the pitch locations for a specific pitcher for several values of the count.

#### Data

The package includes the dataset sc_pitchers_2019 that contains Statcast data for 20 pitchers for the 2019 season.

#### Pitch Locations for a List

Suppose we want to compare the locations of the fastballs thrown by Aaron Nola and Trevor Bauer.  

I find the subset of data I need and then create a list dividing the data by pitcher.

```{r}
d <- filter(sc_pitchers_2019, 
            pitcher %in% c(605400, 545333),
            pitch_type == "FF")
ds <- split(d, d$pitcher)
names(ds) <- c("Bauer", "Nola")
```

Now we can construct the graph.

```{r}
location_compare(ds)
```


#### Pitch Locations for a Specific Count

Suppose we want to look at the locations of Aaron Nola’s pitches on a 0-0 count. I can find Nola’s MLBAM id number by use of the chadwick dataset (also included in the package) that contains the id numbers for all players.


```{r}
chadwick %>% 
  filter(name_last == "Nola", name_first == "Aaron")
```

To produce the graph, type

```{r}
location_count(sc_pitchers_2019, 
               605400, "Aaron Nola", "0-0")
```

#### Pitch Locations Across a Group of Counts

Suppose we want to compare Nola's pitch locations across the counts "0-0", "1-0", "0-1", "0-2"

```{r}
location_count_compare(sc_pitchers_2019, 
               605400, "Aaron Nola", 
               "R", "Offspeed", 
               c("0-0", "1-0", "0-1", "0-2"))
```

