---
title: "Computing Pitch Values"
author: "Jim Albert"
date: "1/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Outline of work on pitch values

This illustrates new functions and data from the CalledStrike package to compute and graph pitch values.

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(CalledStrike)
library(mgcv)
library(metR)
```

#### Read in Statcast data

Read in Statcast data from 2019 season:

```{r}
sc <- read_csv("~/Dropbox/2016 WORK/BLOG Baseball R/OTHER/StatcastData/statcast2019.csv")
```

#### Runs values

The runs value of each count and the runs value for each end-of-PA outcome are contained in the dataset ```count_values``` contained in the ```CalledStrike``` package.

```{r}
count_values
```

#### Compute the pitch values

The function ```compute_pitch_values()``` will create count variables and compute pitch values for all pitches in a Statcast dataset.

```{r}
sc <- compute_pitch_values(sc)
```

#### Remove some plays

Remove plays like stolen bases and pickoff plays (with pitch value of 0) that don't affect pitch value.

```{r}
sc %>%
  filter(abs(Runs.y) > 0) -> sc
```


#### Graph of Smoothed Pitch Values

This gives graph for a specific pitch type.  Note that I have changed the sign on the pitch values so we get perspective from the pitcher.

Pitch values of changeups to right-handed hitters.

```{r}
pitch_value_contour(filter(sc, 
                           pitch_type == "CH",
                           stand == "R"), 
    title = "Pitch Values of Changeups, Stand = R")
```

Compare pitch values of changeups to left and right-handed hitters.

```{r}
changeups <- filter(sc, pitch_type == "CH")
d <- split(changeups, changeups$stand)
pitch_value_contour(d,
      title = "Pitch Value of Changeups")
```

Compare for left and right sided pitchers.

```{r}
sc_L <- filter(sc, p_throws == "L")
sc_R <- filter(sc, p_throws == "R")
d <- c(split(sc_L, sc_L$stand),
      split(sc_R, sc_R$stand))
names(d) <- c("Pitch L, Stand L",
              "Pitch L, Stand R",
              "Pitch R, Stand L",
              "Pitch R, Stand R")
pitch_value_contour(d,
      title = "Pitch Value of Changeups")
```

#### Tile Plots

The function ```pitch_value_plot()``` constructs heat maps using the ```geom_tile()``` geometric object from the ggplot2 package.  They don't require an input of values for the contour lines.

Pitch values of changeups to right-handed hitters.

```{r}
pitch_value_plot(filter(sc, 
                           pitch_type == "CH",
                           stand == "R"), 
    title = "Pitch Values of Changeups, Stand = R")
```










